#**Reproducible Research: Peer Assessment 1**

###**Loading and preprocessing the data**


```{r}

data.file <- "activity.csv"

if (!file.exists(data.file)){
    download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "data.zip", method="curl")
    
  unzip("data.zip")
  file.remove("data.zip")
}


activity <- read.csv(data.file, header= T,sep=",")

activSteps <- activity$steps 
activInterval <- activity$interval  
activDate  <- activity$date <- as.Date(activity$date,"%Y-%m-%d")

```

Print a summary of the data set

```{r}

summary(activity)

```

###**The mean total number of steps taken per day**

Calculate the total number of steps taken per day

```{r}

sumNumStepDay <- tapply(activSteps, activDate, sum)

```

Plot the histogram of the total steps per day

```{r}

hist(sumNumStepDay, col = "green", xlab = "Total Steps per Day", ylab = "Frequency", 
                     main = "Histogram of the total Steps taken per day")

```


Comput the mean of total steps taken per day

```{r}

meanNumStepsDay <- mean(sumNumStepDay,na.rm = TRUE)
meanNumStepsDay

```

Comput the median of total steps taken per day

```{r}

medNumStepsDay  <- median(sumNumStepDay, na.rm = TRUE)
medNumStepsDay
```

###**The average daily activity pattern**

Calculate average of steps in each 5 minutes interval.

```{r}

averNumStepDay <- tapply(activSteps, activInterval, mean, na.rm = TRUE)

```

Timeseries plot of of the 5-minute interval and the average number of steps taken, averaged across all days

```{r}

plot(row.names(averNumStepDay), averNumStepDay, type = "l", xlab = "Time Intervals (5-minute)", 
     ylab = "Mean number of steps taken all days", main = "Timeseries of steps taken at 5 minute Intervals", 
     col = "green")
```

The 5-minute interval on average across all the days in the dataset containing the maximum number of steps

```{r}

maxNumStepInt <- which(averNumStepDay == max(averNumStepDay))
maxAverNumSteps <- max(averNumStepDay)
maxAverNumSteps

```

Find the time interval that contains the maximum average number of steps over all days

```{r}

maxNumStepInt <- which.max(averNumStepDay)
maxAverNumSteps <- names(maxNumStepInt)
maxNumStepInt

```

The `r maxNumStepInt` 5-minute interval contains the 
maximum number of steps on average across all the days

###**Imputing missing values**

Calculate and report the total number of missing values in the dataset

```{r}

idxNaRow  <- is.na(activSteps)
numNaRow <- sum(idxNaRow)
numNaRow

activity$steps[idxNaRow] <- round(mean(activSteps[!idxNaRow]))
```

Fill in missing values using the average interval value across all days

```{r}

naIdx <- which(is.na(activity))
imputNaVal <- averNumStepDay[as.character(activity[naIdx, 3])]
names(imputNaVal) <- naIdx
for (i in naIdx) {
        activity$steps[i] = imputNaVal[as.character(i)]
}

```

Checking that there is no longer NA values in the data

```{r}

sum(is.na(activity))

```

###**Differences in activity patterns between weekdays and weekends**

Creat a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}

weekDays <- weekdays(activDate)

activity$dayType <- ifelse(weekDays == "samedi" | weekDays == "dimanche", "Weekend", 
                           "Weekday")
```

Make a panel plot containing a time series plot 

```{r}
averNumStepDay <- aggregate(activity$steps, by = list(activity$interval, activity$dayType), 
                       mean)

names(averNumStepDay) <- c("interval", "dayType", "steps")

library(lattice)
xyplot(steps ~ interval | dayType, averNumStepDay, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps", col = "green")
```


